\documentclass{article}
\usepackage{graphicx}
\graphicspath{ {./Images/} }
\usepackage[utf8]{inputenc}

\title{STRATEGIES FOR IMPROVING SMALL SCALE ENTERPRISES}
\author{Vibhanshu Singh}
\date{February 2022}

\begin{document}

\maketitle

\section{Abstract}
\section{Introduction}

\newpage

\section*{Abstract}
\newcommand\shortlorem{The study is aimed at the reappraising on the strategies for improving small scale enterprises. The study would help to identify some of the strategies for improving small scale enterprise. It will also help the future investors to be aware of the strategies to be encountered. However, the research question are based on the strategies for improving small scale enterprise, the roles of the government to play in improving small scale enterprises such as providing of basic infrastructural facilities, providing of adequate providing of adequate capital or finance. During the research necessary data was collected through the use of questionnaire and was analysed making use of tabulation method and simple percentage. During the course of the project the survey research method was the research method used. Moreover, the pilot survey was used for the sample size while simple random are used for the sampling technique i.e. thirty five out of forty of the staff were randomly selected. Also effort were made by the Researchers to collect the needed data and accurate information through the use of both primary and secondary sources of data. The data collected was analysed, interpreted and presented using tabulation method and simple percentage. Furthermore the researchers found out that lack of fund, inadequate planning, incompetence of management, poor competitive position are the problems facing small scale enterprises. Also they found out that the government, individual and other organization over the environment have a great role to play in the management of small scale establishment in our country. Finally the researchers recommended that appropriate credit guarantee schemes should be established and the support from credit bureau developed. They also recommended that Development of Infrastructure by the government should be provided to support small scale enterprises.}

\shortlorem



\section*{Introduction}
 


The strategies for improving small scale enterprise in Nigeria is a programme embarked
upon by developing countries as one of the asset of economic increase of their nations.
Before Nigeria had their independence in 1960, the people of Nigeria were engaged in
commerce and in various kinds of occupations like fishing, hunting, buying and selling to
earn their living. These occupations were determined by factors such as geographical
Location and Nature of land, etc.However, they engage on those activities but there were a lot of problems facing them
such as, under-capital, inadequate infrastructure, lack of experience, etc. And these led to
the introduction of a programme known as (SAP) Structural Adjustment programme in
July 1986 by the government of General Ibrahim Badanasi Babangida. This programme
was to help in generating the Nigeria Economy. It went a long way in encouraging private
individuals to participate in the ownership of their own companies.The structural adjustment programme (SAP) Introduced made employment rate to decline
as a result of placing an embargo on employment, retrenchment of workers followed and
many labours both skilled and unskilled could not get employment decided to get a self
employed business such as craft work, drawings and paintings etc.The small scale enterprises is a private owned and operated business characterized by a
small number of employees and low turnover.They are normally private owned co-operations and sole-proprietorship. Small scale
enterprise varies country by country and industry by industry. Taking for instance ina under the work act of 2006; The range is fever than 15 employees. In Europe,
their range is not more than 50employees and fewer than 500 employees to qualify for
U.S small business Administration programs. (Isemin 2009:19) The importance of small scale business enterprise includes the following: small scale Download
enterprise promote competition and hinder monopoly. Small scale enterprise serves as a
nursery for entrepreneurial organization all over the world. They ensure the supply of Pownload
goods and services within the reach of a great majority of the populations. Finally they
employ labour intensive technology and also provides further, the necessary
apprenticeship training in skills and entrepreneurship which enables the beneficiaries get Make Money
self employment. The small scale business which enjoys the patronage of the overall investors were established for the purpose of retailing goods and rendering services Upload Your Academic
required for satisfying the needs and desire of our people in Nigeria. (Norbert Mille). Works and Start Earning
Government Through its various policies has given active encouragement and incentives Cash.
to small scale enterprises in form of banning the inflow of some foreign goods and
sometimes encouraging the importation of certain needed materials and equipment for the
private sector of the economy. In 1987, Government also went further to establish a
programme known as (NDE) National Directorate of Employment in Nigeria aiming at
precisly creating more job opportunities. It also served as a vehicle for promoting
sneurship amona the nation vouths. At inception NDE develoned four core 
programmes which is the youth employment and vocational skills, small scale enterprise
and Graduate employment, special public works, Agricultural and Rural employment
promotion.The youth employment and vocational skills development programme was further split
into four schemes namely:- National open Apprenticeship (NOA) scheme, schools on
wheels (SON) scheme, waste to wealth (WIO) scheme and Resettlement Scheme. Various Download
schemes that were embarked upon three of this programmes in one way or the others
encouraged self employment. They provide their participants with necessary vocational Download
and managerial skills, in order to cope with the demands of their preferred entrepreneurial
activities. They also provide them with part of their initial capital in form of guaranteed
and re-settlement loans as well as tools and equipments. Make Money
Government went on to set up other organizations for the development of small scale
enterprise from where entrepreneurs can obtain the necessary information and other Uefoete| Your Academic
helps they require in order to establish and run their business successfully. These Works and Start TEU]
institutions include: Co-operate Affairs Commission (CAC), Industrial Development Centre Cash.
(IDCS) Centre for Management Development (CMD), National economic Reconstruction
Fund (NEFRUND), Development banks and people bank of Nigeria (DPBN) etc. these
institutions went. Further to promote and develop small scale enterprises throughout the country.




\end{document}